---
title: "frugivore_light"
output: html_notebook
---

# Questions: 
Did bats have a preference for the dark side? (light/ lunarphobia)
 Did habituation have an effect on light phobia?
 Did light experience in habituation have an effect on light phobia?
 
 - Did they EAT DIFFERENTLY, VISIT DIFFERENTLY, or CALL DIFFERENTLY in any of the conditions? 


Secondary questions: 
  - What were the real light levels on the light and dark side? Was it really lighter on the "light side"
  - Were preferences stronger when differences in light levels were stronger?
  - Was there any effect of the moon cycle on choices?
  
 - Did bats make more choices overall when they habituated? 
  
  
If any of the pilot trials are not that different, it would be nice to be able to include them. 


################


# Load packages
--> Add these packages under packages > install if you don't already have them downloaded
```{r}
library(ggplot2)
library(tidyverse)
library(readxl)


```



# import data (You can just use the Environment > "import dataset" > "from excel" tool, and choose your second tab )
```{r}
C_Data <- read_excel("Carollia Project Data.xlsx", 
    sheet = "Trial Info", col_types = c("text", 
        "text", "text", "date", "date", "text", 
        "text", "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))

```

# format and filter data
```{r}

#remove spaces from column names for convenience
colnames(C_Data) <- gsub(" ", "_", colnames(C_Data))
View(C_Data)

#number of rows in data 
nrow(C_Data)

#remove unused treatment
C_Data <- C_Data %>%
      filter(Habituation_Condition != "Dark (2 nights)")  # "!=" means "do not include"
#you can use this expression to remove other bats you need to exclude

#Make Habituation condition an ordered factor for plotting, currently ordered None, Dark, light
C_Data <- C_Data %>%
      mutate(Habituation_Condition = fct_relevel(Habituation_Condition, c("None", "Dark", "Light"))) 
      
```


# explore data, what is in each of the columns? what type of data is each column? 
```{r}
str(C_Data)
```

# make columns with relative number of vists/ food eaten to the dark side (negative = more dark, positive = more light)
```{r}
#columns with dark responses subtracted from light

C_Data <- 
      C_Data %>%
      mutate(rel_Dark_eaten = LIGHT_Total_amount_eaten - DARK_Total_amount_eaten, #relative amount eaten
            rel_Dark_choices = Number_of_choices_at_Light -Number_of_choices_at_Dark, #relative # of visits
            rel_Dark_calls = `Number_of_calls/visit_LIGHT` - `Number_of_calls/visit_DARK` #relative # of calls (you need ` around this variable bc of the "\")
            )
```



#summary of relative preference in the different conditions
```{r}
C_Data %>%
      group_by(Habituation_Condition) %>%
      summarize(mean_rel_eaten = mean(rel_Dark_eaten, na.rm= TRUE), mean_rel_choices = mean(rel_Dark_choices, na.rm = TRUE) )
```


# plot: relative preference to light and dark, by condition. 
y = # visits or amount eaten
x = treatment (light/ dark/ none)
```{r}
# change this to # visits, then subgroup by visits light, visits dark 

# important to know that there are a lot of NAs
C_Data %>%
      ggplot(aes(y= rel_Dark_eaten, x = Habituation_Condition, color = Habituation_Condition)) +
      geom_boxplot() +
       geom_jitter(alpha=0.5) +
      ylab("Relative amt eaten (+ = more in light, - = more in dark)") +
      xlab("Habituation Condition")
#
C_Data %>%
      ggplot(aes(y= rel_Dark_choices, x = Habituation_Condition, color = Habituation_Condition)) +
      geom_boxplot() +
      geom_jitter(alpha=0.5) +
      ylab("Relative number of visits (+ = more to light, - = more to dark)") +
      xlab("Habituation Condition")


C_Data %>%
      ggplot(aes(y= rel_Dark_calls, x = Habituation_Condition, color = Habituation_Condition)) +
      geom_boxplot() +
      geom_jitter(alpha=0.5) +
      ylab("Relative number of calls (+ = more to light, - = more to dark)") +
      xlab("Habituation Condition")

```

##############
Statistical strategies: 
1) lay out questions
2) lay out types of variable: response variable, predictor variable(s) (categorical/ continuous/ )
3) choose possible tests that fit that



1) Classic: ANOVA
2) permutation test (permute the treatment labels.  )
3) general linear model or generalized linear mixed model



